---
title: "SQL-to-R Code"
author: Prithviraj Lakkakula
output: 
  html_document: 
    toc: yes
    highlight: zenburn
    theme: cosmo
  pdf_document: 
    toc: yes
    highlight: zenburn
---

<!--
:::: {style="display: grid;"} ### this code if you add below will display one after the other top and bottom.

#For more info, see here: https://bookdown.org/yihui/rmarkdown-cookbook/multi-column.html
-->

In this post, I will query the data from different data tables in both SQL and R's dplyr with a goal of obtaining the same output. SQL code will be on the your left and R code will be on your right. 

This is a work in progress as I continue to add the queries in the coming weeks starting with simple queries to more complicated queries. 

Note: In the html file, two columns appear side by side but in pdf version you can view the code only one after the other (not side by side).

<!--
## Populating the database (in-memory) with data
-->

```{r echo = FALSE, message=FALSE}
library(readxl)
films <- read_excel("films.xlsx")
people <- read.csv("people.csv")
reviews <- read_excel("reviews.xlsx")
roles <- read_excel("roles.xlsx")

library(odbc)
library(DBI)
library(RSQL)
library(RSQLite)
library(tidyverse)

con <- dbConnect(drv = RSQLite::SQLite(),
                 dbname = ":memory:")

dbWriteTable(conn = con,
             name = "films",
             value = films)
dbWriteTable(conn = con,
             name = "people",
             value = people)
dbWriteTable(conn = con,
             name = "reviews",
             value = reviews)
dbWriteTable(conn = con,
             name = "roles",
             value = roles)
#dbListTables(con)
#dbListFields(con, "people")
```


## Query 1: SELECTing single columns

:::: {style="display: grid; grid-template-columns: 1fr 1fr; grid-column-gap: 10px;"}


::: {}

SQL code

```{sql, connection=con}
SELECT name
FROM people
LIMIT 10;
```

:::

::: {}
R code

```{r warning=FALSE, message=FALSE}
people %>% 
  select(name) %>% 
  head(n = 10)
```
:::

::::


## Query 2: SELECTing multiple columns

:::: {style="display: grid; grid-template-columns: 1fr 1fr; grid-column-gap: 10px;"}


::: {}

SQL code

```{sql, connection=con}
SELECT name, birthdate
FROM people
LIMIT 10;
```

:::

::: {}
R code

```{r warning=FALSE, message=FALSE}
people %>% 
  select(name, birthdate) %>% 
  head(n = 10)
```
:::

::::

## Query 3: SELECTing all columns

:::: {style="display: grid; grid-template-columns: 1fr 1fr; grid-column-gap: 10px;"}


::: {}

SQL code

```{sql, connection=con}
SELECT *
FROM people
LIMIT 10;
```

:::

::: {}
R code

```{r warning=FALSE, message=FALSE}
people %>% 
  head(n = 10)
```
:::

::::

## Query 4: Excluding specific columns

:::: {style="display: grid; grid-template-columns: 1fr 1fr; grid-column-gap: 10px;"}


::: {}

SQL code

```{sql, connection=con}
SELECT id, name, birthdate
FROM people
LIMIT 10;
```

:::

::: {}
R code

```{r warning=FALSE, message=FALSE}
people %>% 
  select(-deathdate) %>%
  head(n = 10)
```
:::

::::

## Query 5: SELECTing DISTINCT columns

:::: {style="display: grid; grid-template-columns: 1fr 1fr; grid-column-gap: 10px;"}


::: {}

SQL code

```{sql, connection=con}
SELECT DISTINCT language
FROM films
LIMIT 10;
```

:::

::: {}
R code

```{r warning=FALSE, message=FALSE}
films %>% 
  distinct(language) %>%
  head(n = 10)
```
:::

::::

:::: {style="display: grid; grid-template-columns: 1fr 1fr; grid-column-gap: 10px;"}


::: {}

SQL code

```{sql, connection=con}
SELECT DISTINCT country
FROM films
LIMIT 10;
```

:::

::: {}
R code

```{r warning=FALSE, message=FALSE}
films %>% 
  distinct(country) %>%
  head(n = 10)
```
:::

::::

:::: {style="display: grid; grid-template-columns: 1fr 1fr; grid-column-gap: 10px;"}


::: {}

SQL code

```{sql, connection=con}
SELECT DISTINCT certification
FROM films
LIMIT 10;
```

:::

::: {}
R code

```{r warning=FALSE, message=FALSE}
films %>% 
  distinct(certification) %>%
  head(n = 10)
```
:::

::::

:::: {style="display: grid; grid-template-columns: 1fr 1fr; grid-column-gap: 10px;"}


::: {}

SQL code

```{sql, connection=con}
SELECT DISTINCT role
FROM roles
LIMIT 10;
```

:::

::: {}
R code

```{r warning=FALSE, message=FALSE}
roles %>% 
  distinct(role) %>%
  head(n = 10)
```
:::

::::

## Query 6: Learning to COUNT

:::: {style="display: grid; grid-template-columns: 1fr 1fr; grid-column-gap: 10px;"}


::: {}

SQL code: Count the number of rows in `people` table

```{sql, connection=con}
SELECT COUNT(*)
FROM people;
```

:::

::: {}
R code: Count the number of rows in `people` table

```{r warning=FALSE, message=FALSE}
people %>% 
  count() 
```
:::

::::

:::: {style="display: grid; grid-template-columns: 1fr 1fr; grid-column-gap: 10px;"}


::: {}

SQL code: Count the number of birth dates in the `people` table

```{sql, connection=con}
SELECT COUNT(birthdate)
FROM people;
```

:::

::: {}
R code: Count the number of birth dates in the `people` table

```{r warning=FALSE, message=FALSE}
people %>% select(birthdate) %>%
  count() 
```
:::

::::

:::: {style="display: grid; grid-template-columns: 1fr 1fr; grid-column-gap: 10px;"}


::: {}

SQL code: Count the number of DISTINCT birth dates in the `people` table

```{sql, connection=con}
SELECT COUNT(DISTINCT birthdate)
FROM people;
```

:::

::: {}
R code: Count the number of DISTINCT birth dates in the `people` table

```{r warning=FALSE, message=FALSE}
people %>% select(birthdate) %>%
  n_distinct()
```
:::

::::

:::: {style="display: grid; grid-template-columns: 1fr 1fr; grid-column-gap: 10px;"}


::: {}

SQL code: Count the number of DISTINCT languages in the `films` table

```{sql, connection=con}
SELECT COUNT(DISTINCT language)
FROM films;
```

:::

::: {}
R code: Count the number of DISTINCT languages in the `films` table

```{r warning=FALSE, message=FALSE}
films %>% select(language) %>%
  n_distinct()
```
:::

::::

::::

:::: {style="display: grid; grid-template-columns: 1fr 1fr; grid-column-gap: 10px;"}


::: {}

SQL code: Count the number of DISTINCT languages in the `films` table

```{sql, connection=con}
SELECT COUNT(DISTINCT country)
FROM films;
```

:::

::: {}
R code: Count the number of DISTINCT languages in the `films` table

```{r warning=FALSE, message=FALSE}
films %>% select(country) %>%
  n_distinct()
```
:::

::::


## Query 7: Filtering of numeric values

:::: {style="display: grid; grid-template-columns: 1fr 1fr; grid-column-gap: 10px;"}


::: {}

SQL code: selects all details for films with a budget over ten thousand dollars

```{sql, connection=con}
SELECT *
FROM films
WHERE budget > 10000
LIMIT 5;
```

:::

::: {}
R code: selects all details for films with a budget over ten thousand dollars

```{r warning=FALSE, message=FALSE}
films %>% 
  filter(budget > 10000) %>%
  head(n = 5)
```
:::

::::


:::: {style="display: grid; grid-template-columns: 1fr 1fr; grid-column-gap: 10px;"}


::: {}

SQL code: selects all details for all films released in 2016

```{sql, connection=con}
SELECT *
FROM films
WHERE release_year = 2016
LIMIT 5;
```

:::

::: {}
R code: selects all details for all films released in 2016

```{r warning=FALSE, message=FALSE}
films %>% 
  filter(release_year == 2016) %>%
  head(n = 5)
```
:::

::::

:::: {style="display: grid; grid-template-columns: 1fr 1fr; grid-column-gap: 10px;"}


::: {}

SQL code: selects number of films released before 2000

```{sql, connection=con}
SELECT COUNT(release_year)
FROM films
WHERE release_year <2000;
```

:::

::: {}
R code: selects number of films released before 2000

```{r warning=FALSE, message=FALSE}
films %>% 
  count(release_year < 2000) 
```
:::

::::

## Query 8: Filtering text

:::: {style="display: grid; grid-template-columns: 1fr 1fr; grid-column-gap: 10px;"}


::: {}

SQL code: gets the titles of all films which were filmed in China

```{sql, connection=con}
SELECT title
FROM films
WHERE country = 'China' -- in PostgreSQL you must use single quotes
LIMIT 5;
```

:::

::: {}
R code: gets the titles of all films which were filmed in China

```{r warning=FALSE, message=FALSE}
films %>% 
  filter(country == "China") %>% # here you must use double quotes around text
  select(title) %>%
  head(n = 5)
```
:::

::::

:::: {style="display: grid; grid-template-columns: 1fr 1fr; grid-column-gap: 10px;"}


::: {}

SQL code: gets all the details for all French language films

```{sql, connection=con}
SELECT *
FROM films
WHERE language = 'French' -- in PostgreSQL you must use single quotes
LIMIT 5;
```

:::

::: {}
R code: gets all the details for all French language films

```{r warning=FALSE, message=FALSE}
films %>% 
  filter(language == "French") %>% # here you must use double quotes around text
  head(n = 5)
```
:::

::::

:::: {style="display: grid; grid-template-columns: 1fr 1fr; grid-column-gap: 10px;"}


::: {}

SQL code: Get the name and birth date of the person born on November 11th, 1974.

```{sql, connection=con}
SELECT name birthdate
FROM people
WHERE birthdate = '1974-11-11' -- in PostgreSQL you must use single quotes
LIMIT 5;
```

:::

::: {}
R code: Get the name and birth date of the person born on November 11th, 1974.

```{r warning=FALSE, message=FALSE}
people %>% 
  select(name, birthdate) %>%
  filter(birthdate == "1974-11-11") %>% # here you must use double quotes around text
  head(n = 5)
```
:::

::::


:::: {style="display: grid; grid-template-columns: 1fr 1fr; grid-column-gap: 10px;"}


::: {}

SQL code: Get the number of Hindi language films

```{sql, connection=con}
SELECT COUNT(language)
FROM films
WHERE language = 'Hindi'; -- in PostgreSQL you must use single quotes
```

:::

::: {}
R code: Get the number of Hindi language films

```{r warning=FALSE, message=FALSE}
films %>% 
  filter(language == "Hindi") %>% # here you must use double quotes around text
  count()
```
:::

::::

## Query 9: Use WHERE and AND for multiple conditions

:::: {style="display: grid; grid-template-columns: 1fr 1fr; grid-column-gap: 10px;"}


::: {}

SQL code: Gets the titles of films released between 1994 and 2000.

```{sql, connection=con}
SELECT title
FROM films
WHERE release_year > 1994
AND release_year < 2000
LIMIT 5;
```

:::

::: {}
R code: Gets the titles of films released between 1994 and 2000.

```{r warning=FALSE, message=FALSE}
films %>% 
  filter(release_year > 1994 &
           release_year < 2000) %>% 
  select(title) %>%
  head(n = 5)
```
:::

::::


:::: {style="display: grid; grid-template-columns: 1fr 1fr; grid-column-gap: 10px;"}


::: {}

SQL code: Get the title and release year for all Spanish language films released before 2000

```{sql, connection=con}
SELECT title, release_year
FROM films
WHERE language = 'Spanish'
AND release_year < 2000;
```

:::

::: {}
R code: Get the title and release year for all Spanish language films released before 2000

```{r warning=FALSE, message=FALSE}
films %>% 
  filter(language ==  "Spanish" &
           release_year < 2000) %>% 
  select(title, release_year)
```
:::

::::


:::: {style="display: grid; grid-template-columns: 1fr 1fr; grid-column-gap: 10px;"}


::: {}

SQL code: Get all details for Spanish language films released after 2000, but before 2010.

```{sql, connection=con}
SELECT *
FROM films
WHERE language = 'Spanish'
AND release_year > 2000
AND release_year < 2010
LIMIT 5;
```

:::

::: {}
R code: Get all details for Spanish language films released after 2000, but before 2010.

```{r warning=FALSE, message=FALSE}
films %>% 
  filter(language ==  "Spanish" &
           release_year > 2000 &
           release_year < 2010) %>%
  head(n = 5)
```
:::

::::

## Query 10: Use WHERE and OR for multiple conditions

:::: {style="display: grid; grid-template-columns: 1fr 1fr; grid-column-gap: 10px;"}


::: {}

SQL code: Gets all films release in either 1994 or 2000

```{sql, connection=con}
SELECT title
FROM films
WHERE release_year = 1994
OR release_year = 2000
LIMIT 5;
```

:::

::: {}
R code: Gets all films release in either 1994 or 2000

```{r warning=FALSE, message=FALSE}
films %>% 
  filter(release_year == 1994 |
           release_year == 2000) %>%
  select(title) %>%
  head(n = 5)
```
:::

::::

## Query 11: Combining AND and OR with WHERE in SQL

:::: {style="display: grid; grid-template-columns: 1fr 1fr; grid-column-gap: 10px;"}


::: {}

SQL code: Gets all films release in either 1994 or 2000

```{sql, connection=con}
SELECT title
FROM films
WHERE (release_year = 1994 OR release_year = 1995)
AND (certification = 'PG' OR certification = 'R')
LIMIT 5;
```

:::

::: {}
R code: Gets all films release in either 1994 or 2000

```{r warning=FALSE, message=FALSE}
films %>% 
  filter((release_year == 1994 | release_year == 1995) &
           (certification == "PG" | certification == "R")) %>%
  select(title) %>%
  head(n = 5)
```
:::

::::


:::: {style="display: grid; grid-template-columns: 1fr 1fr; grid-column-gap: 10px;"}


::: {}

SQL code: Get the title and release year for films released in the 90s.

```{sql, connection=con}
SELECT title, release_year
FROM films
WHERE release_year >= 1994 
AND release_year < 2000
LIMIT 5;
```

:::

::: {}
R code: Get the title and release year for films released in the 90s.

```{r warning=FALSE, message=FALSE}
films %>% 
  filter(release_year >= 1994 & release_year < 2000)  %>%
  select(title, release_year) %>%
  head(n = 5)
```
:::

::::


:::: {style="display: grid; grid-template-columns: 1fr 1fr; grid-column-gap: 10px;"}


::: {}

SQL code: filter the records to only include French or Spanish language films in 1990s.

```{sql, connection=con}
SELECT title, release_year
FROM films
WHERE (release_year >= 1990 AND release_year < 2000)
AND (language = 'French' OR language = 'Spanish')
LIMIT 5;
```

:::

::: {}
R code: filter the records to only include French or Spanish language films in 199Os.

```{r warning=FALSE, message=FALSE}
films %>% 
  filter((release_year >= 1990 & release_year < 2000) &
           (language == "French" | language == "Spanish")) %>%
  select(title, release_year) %>%
  head(n = 5)
```
:::

::::


:::: {style="display: grid; grid-template-columns: 1fr 1fr; grid-column-gap: 10px;"}


::: {}

SQL code: filter the records to only include French or Spanish language films in 1990s with gross greater than 2 million.

```{sql, connection=con}
SELECT title, release_year
FROM films
WHERE (release_year >= 1990 AND release_year < 2000)
AND (language = 'French' OR language = 'Spanish')
AND gross > 2000000
LIMIT 5;
```

:::

::: {}
R code: filter the records to only include French or Spanish language films in 199Os with gross greater than 2 million.

```{r warning=FALSE, message=FALSE}
films %>% 
  filter((release_year >= 1990 & release_year < 2000) &
           (language == "French" | language == "Spanish") &
           gross > 2000000) %>%
  select(title, release_year) %>%
  head(n = 5)
```
:::

::::