---
title: "SQL-to-R Code"
author: Prithviraj Lakkakula
output: 
  html_document: 
    toc: yes
    highlight: zenburn
    theme: cosmo
  pdf_document: 
    toc: yes
    highlight: zenburn
---

<!--
:::: {style="display: grid;"} ### this code if you add below will display one after the other top and bottom.

#For more info, see here: https://bookdown.org/yihui/rmarkdown-cookbook/multi-column.html
-->

In this post, I will query the data from different data tables in both SQL and R's dplyr with a goal of obtaining the same output. SQL code will be on the your left and R code will be on your right. 

This is a work in progress as I continue to add the queries in the coming weeks starting with simple queries to more complicated queries. 

Note: In the html file, two columns appear side by side but in pdf version you can view the code only one after the other (not side by side).

### Query No. 1

:::: {style="display: flex;"}

::: {}
Here is the **first** Div.

```{r}
## SQL Code will come here
```
:::

::: {}
And this block will be put on the right:

```{r}
## R code will go here
```
:::

::::

<!--
### Query No. 2

:::: {style="display: flex;"}

::: {}
Here is the **first** Div.

```{r}
str(iris)
```
:::

::: {}
And this block will be put on the right:

```{r}
plot(iris[, -5])
```
:::

::::
-->

