---
title: "SQL-to-R Code"
author: Prithviraj Lakkakula
output: 
  html_document: 
    toc: yes
    highlight: zenburn
    theme: cosmo
  pdf_document: 
    toc: yes
    highlight: zenburn
---

```{r, include=FALSE}
options(tinytex.verbose = TRUE)
```
<!--
:::: {style="display: grid;"} ### this code if you add below will display one after the other top and bottom.

#For more info, see here: https://bookdown.org/yihui/rmarkdown-cookbook/multi-column.html
-->

In this post, I will query the data from different data tables in both SQL and R's dplyr with a goal of obtaining the same output. SQL code will be on the your left and R code will be on your right. 

This is a work in progress as I continue to add the queries in the coming weeks starting with simple queries to more complicated queries. 

Note: In the html file, two columns appear side by side but in pdf version you can view the code only one after the other (not side by side).

<!--
## Populating the database (in-memory) with data
-->

```{r echo = FALSE, message=FALSE}
library(readxl)
films <- read_excel("films.xlsx")
people <- read.csv("people.csv")
reviews <- read_excel("reviews.xlsx")
roles <- read_excel("roles.xlsx")
#joining data in sql - data
cities <- read_excel("cities.xlsx")
countries <- read_excel("countries.xlsx")
economies <- read_excel("economies.xlsx")
languages <- read_excel("languages.xlsx")
populations <- read_excel("populations.xlsx")

library(odbc)
library(DBI)
library(RSQL)
library(RSQLite)
library(tidyverse)
library(data.table)
library(reprex)

con <- dbConnect(drv = RSQLite::SQLite(),
                 dbname = ":memory:")

dbWriteTable(conn = con,
             name = "films",
             value = films)
dbWriteTable(conn = con,
             name = "people",
             value = people)
dbWriteTable(conn = con,
             name = "reviews",
             value = reviews)
dbWriteTable(conn = con,
             name = "roles",
             value = roles)
dbWriteTable(conn = con,
             name = "cities",
             value = cities)
dbWriteTable(conn = con,
             name = "countries",
             value = countries)
dbWriteTable(conn = con,
             name = "economies",
             value = economies)
dbWriteTable(conn = con,
             name = "languages",
             value = languages)
dbWriteTable(conn = con,
             name = "populations",
             value = populations)
#dbListTables(con)
#dbListFields(con, "people")
```


# Introduction to SQL

## Query 1: SELECTing single columns

:::: {style="display: grid; grid-template-columns: 1fr 1fr; grid-column-gap: 10px;"}


::: {}

SQL code

```{sql, connection=con}
SELECT name
FROM people
LIMIT 10;
```

:::

::: {}
R code

```{r warning=FALSE, message=FALSE}
people %>% 
  select(name) %>% 
  head(n = 10)
```
:::

::::


## Query 2: SELECTing multiple columns

:::: {style="display: grid; grid-template-columns: 1fr 1fr; grid-column-gap: 10px;"}


::: {}

SQL code

```{sql, connection=con}
SELECT name, birthdate
FROM people
LIMIT 10;
```

:::

::: {}
R code

```{r warning=FALSE, message=FALSE}
people %>% 
  select(name, birthdate) %>% 
  head(n = 10)
```
:::

::::

## Query 3: SELECTing all columns

:::: {style="display: grid; grid-template-columns: 1fr 1fr; grid-column-gap: 10px;"}


::: {}

SQL code

```{sql, connection=con}
SELECT *
FROM people
LIMIT 10;
```

:::

::: {}
R code

```{r warning=FALSE, message=FALSE}
people %>% 
  head(n = 10)
```
:::

::::

## Query 4: Excluding specific columns

:::: {style="display: grid; grid-template-columns: 1fr 1fr; grid-column-gap: 10px;"}


::: {}

SQL code

```{sql, connection=con}
SELECT id, name, birthdate
FROM people
LIMIT 10;
```

:::

::: {}
R code

```{r warning=FALSE, message=FALSE}
people %>% 
  select(-deathdate) %>%
  head(n = 10)
```
:::

::::

## Query 5: SELECTing DISTINCT columns

:::: {style="display: grid; grid-template-columns: 1fr 1fr; grid-column-gap: 10px;"}


::: {}

SQL code

```{sql, connection=con}
SELECT DISTINCT language
FROM films
LIMIT 10;
```

:::

::: {}
R code

```{r warning=FALSE, message=FALSE}
films %>% 
  distinct(language) %>%
  head(n = 10)
```
:::

::::

:::: {style="display: grid; grid-template-columns: 1fr 1fr; grid-column-gap: 10px;"}


::: {}

SQL code

```{sql, connection=con}
SELECT DISTINCT country
FROM films
LIMIT 10;
```

:::

::: {}
R code

```{r warning=FALSE, message=FALSE}
films %>% 
  distinct(country) %>%
  head(n = 10)
```
:::

::::

:::: {style="display: grid; grid-template-columns: 1fr 1fr; grid-column-gap: 10px;"}


::: {}

SQL code

```{sql, connection=con}
SELECT DISTINCT certification
FROM films
LIMIT 10;
```

:::

::: {}
R code

```{r warning=FALSE, message=FALSE}
films %>% 
  distinct(certification) %>%
  head(n = 10)
```
:::

::::

:::: {style="display: grid; grid-template-columns: 1fr 1fr; grid-column-gap: 10px;"}


::: {}

SQL code

```{sql, connection=con}
SELECT DISTINCT role
FROM roles
LIMIT 10;
```

:::

::: {}
R code

```{r warning=FALSE, message=FALSE}
roles %>% 
  distinct(role) %>%
  head(n = 10)
```
:::

::::

## Query 6: Learning to COUNT

:::: {style="display: grid; grid-template-columns: 1fr 1fr; grid-column-gap: 10px;"}


::: {}

SQL code: Count the number of rows in `people` table

```{sql, connection=con}
SELECT COUNT(*)
FROM people;
```

:::

::: {}
R code: Count the number of rows in `people` table

```{r warning=FALSE, message=FALSE}
people %>% 
  count() 
```
:::

::::

:::: {style="display: grid; grid-template-columns: 1fr 1fr; grid-column-gap: 10px;"}


::: {}

SQL code: Count the number of birth dates in the `people` table

```{sql, connection=con}
SELECT COUNT(birthdate)
FROM people;
```

:::

::: {}
R code: Count the number of birth dates in the `people` table

```{r warning=FALSE, message=FALSE}
people %>% select(birthdate) %>%
  count() 
```
:::

::::

:::: {style="display: grid; grid-template-columns: 1fr 1fr; grid-column-gap: 10px;"}


::: {}

SQL code: Count the number of DISTINCT birth dates in the `people` table

```{sql, connection=con}
SELECT COUNT(DISTINCT birthdate)
FROM people;
```

:::

::: {}
R code: Count the number of DISTINCT birth dates in the `people` table

```{r warning=FALSE, message=FALSE}
people %>% select(birthdate) %>%
  n_distinct()
```
:::

::::

:::: {style="display: grid; grid-template-columns: 1fr 1fr; grid-column-gap: 10px;"}


::: {}

SQL code: Count the number of DISTINCT languages in the `films` table

```{sql, connection=con}
SELECT COUNT(DISTINCT language)
FROM films;
```

:::

::: {}
R code: Count the number of DISTINCT languages in the `films` table

```{r warning=FALSE, message=FALSE}
films %>% select(language) %>%
  n_distinct()
```
:::

::::

::::

:::: {style="display: grid; grid-template-columns: 1fr 1fr; grid-column-gap: 10px;"}


::: {}

SQL code: Count the number of DISTINCT languages in the `films` table

```{sql, connection=con}
SELECT COUNT(DISTINCT country)
FROM films;
```

:::

::: {}
R code: Count the number of DISTINCT languages in the `films` table

```{r warning=FALSE, message=FALSE}
films %>% select(country) %>%
  n_distinct()
```
:::

::::


## Query 7: Filtering of numeric values

:::: {style="display: grid; grid-template-columns: 1fr 1fr; grid-column-gap: 10px;"}


::: {}

SQL code: selects all details for films with a budget over ten thousand dollars

```{sql, connection=con}
SELECT *
FROM films
WHERE budget > 10000
LIMIT 5;
```

:::

::: {}
R code: selects all details for films with a budget over ten thousand dollars

```{r warning=FALSE, message=FALSE}
films %>% 
  filter(budget > 10000) %>%
  head(n = 5)
```
:::

::::


:::: {style="display: grid; grid-template-columns: 1fr 1fr; grid-column-gap: 10px;"}


::: {}

SQL code: selects all details for all films released in 2016

```{sql, connection=con}
SELECT *
FROM films
WHERE release_year = 2016
LIMIT 5;
```

:::

::: {}
R code: selects all details for all films released in 2016

```{r warning=FALSE, message=FALSE}
films %>% 
  filter(release_year == 2016) %>%
  head(n = 5)
```
:::

::::

:::: {style="display: grid; grid-template-columns: 1fr 1fr; grid-column-gap: 10px;"}


::: {}

SQL code: selects number of films released before 2000

```{sql, connection=con}
SELECT COUNT(release_year)
FROM films
WHERE release_year <2000;
```

:::

::: {}
R code: selects number of films released before 2000

```{r warning=FALSE, message=FALSE}
films %>% 
  count(release_year < 2000) 
```
:::

::::

## Query 8: Filtering text

:::: {style="display: grid; grid-template-columns: 1fr 1fr; grid-column-gap: 10px;"}


::: {}

SQL code: gets the titles of all films which were filmed in China

```{sql, connection=con}
SELECT title
FROM films
WHERE country = 'China' -- in PostgreSQL you must use single quotes
LIMIT 5;
```

:::

::: {}
R code: gets the titles of all films which were filmed in China

```{r warning=FALSE, message=FALSE}
films %>% 
  filter(country == "China") %>% # here you must use double quotes around text
  select(title) %>%
  head(n = 5)
```
:::

::::

:::: {style="display: grid; grid-template-columns: 1fr 1fr; grid-column-gap: 10px;"}


::: {}

SQL code: gets all the details for all French language films

```{sql, connection=con}
SELECT *
FROM films
WHERE language = 'French' -- in PostgreSQL you must use single quotes
LIMIT 5;
```

:::

::: {}
R code: gets all the details for all French language films

```{r warning=FALSE, message=FALSE}
films %>% 
  filter(language == "French") %>% # here you must use double quotes around text
  head(n = 5)
```
:::

::::

:::: {style="display: grid; grid-template-columns: 1fr 1fr; grid-column-gap: 10px;"}


::: {}

SQL code: Get the name and birth date of the person born on November 11th, 1974.

```{sql, connection=con}
SELECT name birthdate
FROM people
WHERE birthdate = '1974-11-11' -- in PostgreSQL you must use single quotes
LIMIT 5;
```

:::

::: {}
R code: Get the name and birth date of the person born on November 11th, 1974.

```{r warning=FALSE, message=FALSE}
people %>% 
  select(name, birthdate) %>%
  filter(birthdate == "1974-11-11") %>% # here you must use double quotes around text
  head(n = 5)
```
:::

::::


:::: {style="display: grid; grid-template-columns: 1fr 1fr; grid-column-gap: 10px;"}


::: {}

SQL code: Get the number of Hindi language films

```{sql, connection=con}
SELECT COUNT(language)
FROM films
WHERE language = 'Hindi'; -- in PostgreSQL you must use single quotes
```

:::

::: {}
R code: Get the number of Hindi language films

```{r warning=FALSE, message=FALSE}
films %>% 
  filter(language == "Hindi") %>% # here you must use double quotes around text
  count()
```
:::

::::

## Query 9: Use WHERE and AND for multiple conditions

:::: {style="display: grid; grid-template-columns: 1fr 1fr; grid-column-gap: 10px;"}


::: {}

SQL code: Gets the titles of films released between 1994 and 2000.

```{sql, connection=con}
SELECT title
FROM films
WHERE release_year > 1994
AND release_year < 2000
LIMIT 5;
```

:::

::: {}
R code: Gets the titles of films released between 1994 and 2000.

```{r warning=FALSE, message=FALSE}
films %>% 
  filter(release_year > 1994 &
           release_year < 2000) %>% 
  select(title) %>%
  head(n = 5)
```
:::

::::


:::: {style="display: grid; grid-template-columns: 1fr 1fr; grid-column-gap: 10px;"}


::: {}

SQL code: Get the title and release year for all Spanish language films released before 2000

```{sql, connection=con}
SELECT title, release_year
FROM films
WHERE language = 'Spanish'
AND release_year < 2000;
```

:::

::: {}
R code: Get the title and release year for all Spanish language films released before 2000

```{r warning=FALSE, message=FALSE}
films %>% 
  filter(language ==  "Spanish" &
           release_year < 2000) %>% 
  select(title, release_year)
```
:::

::::


:::: {style="display: grid; grid-template-columns: 1fr 1fr; grid-column-gap: 10px;"}


::: {}

SQL code: Get all details for Spanish language films released after 2000, but before 2010.

```{sql, connection=con}
SELECT *
FROM films
WHERE language = 'Spanish'
AND release_year > 2000
AND release_year < 2010
LIMIT 5;
```

:::

::: {}
R code: Get all details for Spanish language films released after 2000, but before 2010.

```{r warning=FALSE, message=FALSE}
films %>% 
  filter(language ==  "Spanish" &
           release_year > 2000 &
           release_year < 2010) %>%
  head(n = 5)
```
:::

::::

## Query 10: Use WHERE and OR for multiple conditions

:::: {style="display: grid; grid-template-columns: 1fr 1fr; grid-column-gap: 10px;"}


::: {}

SQL code: Gets all films release in either 1994 or 2000

```{sql, connection=con}
SELECT title
FROM films
WHERE release_year = 1994
OR release_year = 2000
LIMIT 5;
```

:::

::: {}
R code: Gets all films release in either 1994 or 2000

```{r warning=FALSE, message=FALSE}
films %>% 
  filter(release_year == 1994 |
           release_year == 2000) %>%
  select(title) %>%
  head(n = 5)
```
:::

::::

## Query 11: Combining AND and OR with WHERE in SQL

:::: {style="display: grid; grid-template-columns: 1fr 1fr; grid-column-gap: 10px;"}


::: {}

SQL code: Gets all films release in either 1994 or 2000

```{sql, connection=con}
SELECT title
FROM films
WHERE (release_year = 1994 OR release_year = 1995)
AND (certification = 'PG' OR certification = 'R')
LIMIT 5;
```

:::

::: {}
R code: Gets all films release in either 1994 or 2000

```{r warning=FALSE, message=FALSE}
films %>% 
  filter((release_year == 1994 | release_year == 1995) &
           (certification == "PG" | certification == "R")) %>%
  select(title) %>%
  head(n = 5)
```
:::

::::


:::: {style="display: grid; grid-template-columns: 1fr 1fr; grid-column-gap: 10px;"}


::: {}

SQL code: Get the title and release year for films released in the 90s.

```{sql, connection=con}
SELECT title, release_year
FROM films
WHERE release_year >= 1994 
AND release_year < 2000
LIMIT 5;
```

:::

::: {}
R code: Get the title and release year for films released in the 90s.

```{r warning=FALSE, message=FALSE}
films %>% 
  filter(release_year >= 1994 & release_year < 2000)  %>%
  select(title, release_year) %>%
  head(n = 5)
```
:::

::::


:::: {style="display: grid; grid-template-columns: 1fr 1fr; grid-column-gap: 10px;"}


::: {}

SQL code: filter the records to only include French or Spanish language films in 1990s.

```{sql, connection=con}
SELECT title, release_year
FROM films
WHERE (release_year >= 1990 AND release_year < 2000)
AND (language = 'French' OR language = 'Spanish')
LIMIT 5;
```

:::

::: {}
R code: filter the records to only include French or Spanish language films in 199Os.

```{r warning=FALSE, message=FALSE}
films %>% 
  filter((release_year >= 1990 & release_year < 2000) &
           (language == "French" | language == "Spanish")) %>%
  select(title, release_year) %>%
  head(n = 5)
```
:::

::::


:::: {style="display: grid; grid-template-columns: 1fr 1fr; grid-column-gap: 10px;"}


::: {}

SQL code: filter the records to only include French or Spanish language films in 1990s with gross greater than 2 million.

```{sql, connection=con}
SELECT title, release_year
FROM films
WHERE (release_year >= 1990 AND release_year < 2000)
AND (language = 'French' OR language = 'Spanish')
AND gross > 2000000
LIMIT 5;
```

:::

::: {}
R code: filter the records to only include French or Spanish language films in 199Os with gross greater than 2 million.

```{r warning=FALSE, message=FALSE}
films %>% 
  filter((release_year >= 1990 & release_year < 2000) &
           (language == "French" | language == "Spanish") &
           gross > 2000000) %>%
  select(title, release_year) %>%
  head(n = 5)
```
:::

::::

## Query 12: `BETWEEN` with `AND` and `OR` 


:::: {style="display: grid; grid-template-columns: 1fr 1fr; grid-column-gap: 10px;"}


::: {}

SQL code: `BETWEEN` is used to filter values in a specified range and it is always inclusive

```{sql, connection=con}
SELECT title, release_year
FROM films
WHERE release_year
BETWEEN 1994 AND 2000
LIMIT 5;
```

:::

::: {}
R code: `between` is used in filter our values based on a range of values

```{r warning=FALSE, message=FALSE}
films %>% 
  filter(between(release_year, 1994, 2000))%>%
  select(title, release_year) %>%
  head(n = 5)
```
:::

::::


:::: {style="display: grid; grid-template-columns: 1fr 1fr; grid-column-gap: 10px;"}


::: {}

SQL code: `BETWEEN` when used with both `AND` and `OR`

```{sql, connection=con}
SELECT title, release_year
FROM films
WHERE release_year BETWEEN 1990 AND 2000
AND language = 'Spanish'
AND budget > 100000000
LIMIT 5;
```

:::

::: {}
R code: 

```{r warning=FALSE, message=FALSE}
films %>% 
  filter(between(release_year, 1990, 2000) &
           language == "Spanish" &
           budget > 100000000)%>%
  select(title, release_year) %>%
  head(n = 5)
```
:::

::::

## Query 13: `WHERE IN`

:::: {style="display: grid; grid-template-columns: 1fr 1fr; grid-column-gap: 10px;"}


::: {}

SQL code: If you have many either or conditions on a column, we need to specify several `OR` conditions using `WHERE`. Instead we can used `WHERE IN`.

```{sql, connection=con}
SELECT title, release_year
FROM films
WHERE release_year IN (1990, 2000)
AND duration > 120
LIMIT 5;
```

:::

::: {}
R code: 

```{r warning=FALSE, message=FALSE}
films %>% 
  filter(release_year %in% c(1990, 2000) &
           duration > 120) %>%
  select(title, release_year) %>%
  head(n = 5)
```
:::

::::


:::: {style="display: grid; grid-template-columns: 1fr 1fr; grid-column-gap: 10px;"}


::: {}

SQL code: 

```{sql, connection=con}
SELECT title, release_year
FROM films
WHERE language IN ('English', 'Spanish', 'French')
LIMIT 5;
```

:::

::: {}
R code: 

```{r warning=FALSE, message=FALSE}
films %>% 
  filter(language %in% c("English", "Spanish", "French")) %>%
  select(title, release_year) %>%
  head(n = 5)
```
:::

::::


:::: {style="display: grid; grid-template-columns: 1fr 1fr; grid-column-gap: 10px;"}


::: {}

SQL code: 

```{sql, connection=con}
SELECT title, certification
FROM films
WHERE certification IN ('NC-17', 'R')
LIMIT 5;
```

:::

::: {}
R code: 

```{r warning=FALSE, message=FALSE}
films %>% 
  filter(certification %in% c("NC-17", "R")) %>%
  select(title, certification) %>%
  head(n = 5)
```
:::

::::


## Query 14: Intro to `NULL` and `IS NULL`

:::: {style="display: grid; grid-template-columns: 1fr 1fr; grid-column-gap: 10px;"}


::: {}

SQL code: In SQL, `NULL` represents a missing or unknown value. One can check `NULL` values using `IS NULL`. For example, to count the number of missing birth dates in the people table;

```{sql, connection=con}
SELECT COUNT(*)
FROM people
WHERE birthdate IS NULL
LIMIT 5;
```

:::

::: {}
R code: 

```{r warning=FALSE, message=FALSE}
people %>% 
  filter(is.na(birthdate)) %>% count() %>%
  head(n = 5)
```
:::

::::


:::: {style="display: grid; grid-template-columns: 1fr 1fr; grid-column-gap: 10px;"}


::: {}

SQL code: Sometimes, you will want to filter out missing values so you only get results which are not NULL. To do this, you can use the IS NOT NULL operator.

```{sql, connection=con}
SELECT name
FROM people
WHERE birthdate IS NOT NULL
LIMIT 5;
```

:::

::: {}
R code: 

```{r warning=FALSE, message=FALSE}
people %>% 
  filter(!is.na(birthdate)) %>% 
  select(name) %>%
  head(n = 5)
```
:::

::::


:::: {style="display: grid; grid-template-columns: 1fr 1fr; grid-column-gap: 10px;"}


::: {}

SQL code: Get the number of films which don't have a language associated with them

```{sql, connection=con}
SELECT COUNT(*)
FROM films
WHERE language IS NULL
LIMIT 5;
```

:::

::: {}
R code: 

```{r warning=FALSE, message=FALSE}
films %>% 
  filter(is.na(language)) %>% count() %>%
  head(n = 5)
```
:::

::::


## Query 15: Aggregate Functions

Essentially SQL has some aggregate functions that you can perform on your variables.

:::: {style="display: grid; grid-template-columns: 1fr 1fr; grid-column-gap: 10px;"}


::: {}

SQL code: The following query gives the average value from the budget column of the films table.

```{sql, connection=con}
SELECT AVG(budget) AS budget_ave
FROM films
LIMIT 5;
```

:::

::: {}
R code: 

```{r warning=FALSE, message=FALSE}
films %>% 
  summarize(budget_avg = mean(budget, na.rm = TRUE)) %>%
  head(n = 5)
```
:::

::::


:::: {style="display: grid; grid-template-columns: 1fr 1fr; grid-column-gap: 10px;"}


::: {}

SQL code: The following query gives the maximum value from the budget column of the films table.

```{sql, connection=con}
SELECT MAX(budget) AS budget_max
FROM films
LIMIT 5;
```

:::

::: {}
R code: 

```{r warning=FALSE, message=FALSE}
films %>% 
  summarize(budget_max = max(budget, na.rm = TRUE)) %>%
  head(n = 5)
```
:::

::::


:::: {style="display: grid; grid-template-columns: 1fr 1fr; grid-column-gap: 10px;"}


::: {}

SQL code: The following query gives the minimum value from the budget column of the films table.

```{sql, connection=con}
SELECT MIN(budget) AS budget_min
FROM films
LIMIT 5;
```

:::

::: {}
R code: 

```{r warning=FALSE, message=FALSE}
films %>% 
  summarize(budget_min = min(budget, na.rm = TRUE)) %>%
  head(n = 5)
```
:::

::::


:::: {style="display: grid; grid-template-columns: 1fr 1fr; grid-column-gap: 10px;"}


::: {}

SQL code: The following query gives the minimum value from the budget column of the films table.

```{sql, connection=con}
SELECT SUM(budget) AS budget_sum
FROM films
LIMIT 5;
```

:::

::: {}
R code: 

```{r warning=FALSE, message=FALSE}
films %>% 
  summarize(budget_sum = sum(budget, na.rm = TRUE)) %>%
  head(n = 5)
```
:::

::::


## Query 16: Combining Aggregate functions with WHERE

:::: {style="display: grid; grid-template-columns: 1fr 1fr; grid-column-gap: 10px;"}


::: {}

SQL code: Gets the total budget of movies made in the year 2010 or later.

```{sql, connection=con}
SELECT SUM(budget) 
FROM films
WHERE release_year >=2010
LIMIT 5;
```

:::

::: {}
R code: 

```{r warning=FALSE, message=FALSE}
films %>% 
  filter(release_year >= 2010) %>%
  summarize(budget_sum = sum(budget, na.rm = TRUE)) %>%
  head(n = 5)
```
:::

::::


:::: {style="display: grid; grid-template-columns: 1fr 1fr; grid-column-gap: 10px;"}


::: {}

SQL code: Get the amount grossed by the best performing film between 2000 and 2012, inclusive.

```{sql, connection=con}
SELECT MAX(gross) 
FROM films
WHERE release_year BETWEEN 2000 AND 2012
LIMIT 5;
```

:::

::: {}
R code: 

```{r warning=FALSE, message=FALSE}
films %>% 
  filter(between(release_year, 2000, 2012)) %>%
  summarize(gross_max = max(gross, na.rm = TRUE)) %>%
  head(n = 5)
```
:::

::::

<!--
## Query 17: A Note on Arithmetic

In addition to using aggregate functions, you can perform basic arithmetic with symbols like `+, -, *, and /`.

:::: {style="display: grid; grid-template-columns: 1fr 1fr; grid-column-gap: 10px;"}


::: {}

SQL code: The following gives a result of 12

```{sql, connection=con}
SELECT (4 * 3)
```

:::

::: {}
R code: 

```{r warning=FALSE, message=FALSE}
(4 * 3)
```
:::

::::


:::: {style="display: grid; grid-template-columns: 1fr 1fr; grid-column-gap: 10px;"}


::: {}

SQL code: However, the following query gives a result of 1. This is because SQL assumes that if you divide an integer by an integer, you want to get an integer back. So, we need to be very careful while dividing in SQL. If you want more precision when dividing, you can add decimal places to your numbers

```{sql, connection=con}
SELECT (4 / 3)
```

:::

::: {}
R code: 

```{r warning=FALSE, message=FALSE}
as.integer(4 / 3)
```
:::

::::


:::: {style="display: grid; grid-template-columns: 1fr 1fr; grid-column-gap: 10px;"}


::: {}

SQL code: If you want more precision when dividing, you can add decimal places to your numbers

```{sql, connection=con}
SELECT (4.0 / 3.0)
```

:::

::: {}
R code: 

```{r warning=FALSE, message=FALSE}
(4 / 3)
```
:::

::::
-->

:::: {style="display: grid; grid-template-columns: 1fr 1fr; grid-column-gap: 10px;"}


::: {}

SQL code: Get the number of decades the films table covers. Alias the result as number_of_decades. 

```{sql, connection=con}
SELECT (MAX(release_year) - MIN(release_year)) / 10 AS number_of_decades
FROM films
LIMIT 5;
```

:::

::: {}
R code: 

```{r warning=FALSE, message=FALSE}
films %>%
  summarize(number_of_decades = (max(release_year, na.rm = TRUE) - 
                                   min(release_year, na.rm = TRUE)) / 10) %>%
  head(n = 5)
```
:::

::::


## Query 17: Sorting and Grouping


:::: {style="display: grid; grid-template-columns: 1fr 1fr; grid-column-gap: 10px;"}


::: {}

SQL code: In SQL, the `ORDER BY` keyword is used to sort results in ascending or descending order according to the values of one or more columns. By default `ORDER BY` will sort in ascending order. In order to sort the results in descending order, you can use `DESC` order. 

For example, the following query gives you the titles of films sorted by release year, from newest to oldest.

```{sql, connection=con}
SELECT title
FROM films
ORDER BY release_year DESC
LIMIT 5;
```

:::

::: {}
R code: Analogous verb in R for sorting is `arrange`. For `arrange` the default is ascending but for getting descending order `desc` can be used like in SQL.

```{r warning=FALSE, message=FALSE}
films %>%
  arrange(desc(release_year)) %>%
  select(title) %>%
  head(n = 5)
```
:::

::::


:::: {style="display: grid; grid-template-columns: 1fr 1fr; grid-column-gap: 10px;"}


::: {}

SQL code: For text values, ORDER BY sorts a column by default alphabetically (A to Z)

```{sql, connection=con}
SELECT name
FROM people
ORDER BY name
LIMIT 5;
```

:::

::: {}
R code: 

```{r warning=FALSE, message=FALSE}
people %>%
  arrange(name) %>%
  select(name) %>%
  head(n = 5)
```
:::

::::


:::: {style="display: grid; grid-template-columns: 1fr 1fr; grid-column-gap: 10px;"}


::: {}

SQL code: Get the title of films released in 2000 or 2012, in the order they were released.

```{sql, connection=con}
SELECT title
FROM films
WHERE release_year IN (2000, 2012)
ORDER BY release_year
LIMIT 5;
```

:::

::: {}
R code: Get the title of films released in 2000 or 2012, in the order they were released.

```{r warning=FALSE, message=FALSE}
films %>%
  filter(release_year %in% c(2000, 2012)) %>%
  arrange(release_year) %>%
  select(title) %>%
  head(n = 5)
```
:::

::::


:::: {style="display: grid; grid-template-columns: 1fr 1fr; grid-column-gap: 10px;"}


::: {}

SQL code: Get all details for all films except those released in 2015 and order them by duration.

```{sql, connection=con}
SELECT *
FROM films
WHERE release_year < 2015
OR release_year > 2015
ORDER BY duration
LIMIT 5;
```

:::

::: {}
R code: Get all details for all films except those released in 2015 and order them by duration.

```{r warning=FALSE, message=FALSE}
films %>%
  filter(release_year < 2015 | release_year > 2015) %>%
  arrange(duration) %>%
  head(n = 5)
```
:::

::::


## Query 18: SORTING single columns by DESC

To order results in descending order, you can put the keyword `DESC` after your `ORDER BY`. 

:::: {style="display: grid; grid-template-columns: 1fr 1fr; grid-column-gap: 10px;"}


::: {}

SQL code: Gets all the names in the people table, in reverse alphabetical order.

```{sql, connection=con}
SELECT name
FROM people
ORDER BY name DESC
LIMIT 5;
```

:::

::: {}
R code: Gets all the names in the people table, in reverse alphabetical order.

```{r warning=FALSE, message=FALSE}
people %>%
  arrange(desc(name)) %>%
  select(name) %>%
  head(n = 5)
```
:::

::::


## Query 19: SORTING multiple columns

ORDER BY can also be used to sort on multiple columns. It will sort by the first column specified, then sort by the next, then the next, and so on. For example, the following query sorts birth dates first (oldest to newest) and then sorts on the names in alphabetical order. THE ORDER OF COLUMNS IS IMPORTANT!

Note: To specify multiple columns you separate the column names with a comma. 

:::: {style="display: grid; grid-template-columns: 1fr 1fr; grid-column-gap: 10px;"}


::: {}

SQL code: Gets all the names in the people table, in reverse alphabetical order.

```{sql, connection=con}
SELECT birthdate, name
FROM people
ORDER BY birthdate, name
LIMIT 5;
```

:::

::: {}
R code: Gets all the names in the people table, in reverse alphabetical order.

```{r warning=FALSE, message=FALSE}
people %>%
  arrange(birthdate, name) %>%
  select(birthdate, name) %>%
  head(n = 5)
```
:::

::::


:::: {style="display: grid; grid-template-columns: 1fr 1fr; grid-column-gap: 10px;"}


::: {}

SQL code: Get the `names` and `birthdates` of people ordered by name and birth date. Notice how the second column you order on only steps in when the first column is not decisive to tell the order. The second column acts as a tie breaker

```{sql, connection=con}
SELECT name, birthdate
FROM people
ORDER BY name, birthdate
LIMIT 5;
```

:::

::: {}
R code: Get the `names` and `birthdates` of people ordered by name and birth date. Notice how the second column you order on only steps in when the first column is not decisive to tell the order. The second column acts as a tie breaker

```{r warning=FALSE, message=FALSE}
people %>%
  arrange(name, birthdate) %>%
  select(name, birthdate) %>%
  head(n = 5)
```
:::

::::


## Query 20: GROUP BY

- We saw `ORDER BY` for sorting. Here we use `GROUP BY` for aggregating results.
- Commonly, `GROUP BY` is used with aggregate functions like `COUNT()` or `MAX()`. 
- Note that `GROUP BY` always goes after the `FROM` clause!
- Overall, `GROUP BY` is performing operations within groups.
- Warning: SQL will return an error if you try to `SELECT` a field that is not in your `GROUP BY` clause without using it to calculate some kind of value about the entire group.
- Note that you can combine `GROUP BY` with `ORDER BY` to group your results, calculate something about them, and then order your results. 
- **NOTE**: `ORDER BY` always goes after `GROUP BY`

:::: {style="display: grid; grid-template-columns: 1fr 1fr; grid-column-gap: 10px;"}


::: {}

SQL code: Gets the release year and count of films released in each year

```{sql, connection=con}
SELECT release_year, COUNT(*)
FROM films
GROUP BY release_year
LIMIT 5;
```

:::

::: {}
R code: Gets the release year and count of films released in each year

```{r warning=FALSE, message=FALSE}
films %>%
  group_by(release_year) %>% 
  select(release_year) %>% count() %>%
  head(n = 5)
```
:::

::::


:::: {style="display: grid; grid-template-columns: 1fr 1fr; grid-column-gap: 10px;"}


::: {}

SQL code: Get the release year and average duration of all films, grouped by release year

```{sql, connection=con}
SELECT release_year, AVG(duration)
FROM films
GROUP BY release_year
LIMIT 5;
```

:::

::: {}
R code: Get the release year and average duration of all films, grouped by release year

```{r warning=FALSE, message=FALSE}
films %>%
  group_by(release_year) %>% summarize(duration_avg = mean(duration)) %>%
  select(release_year, duration_avg) %>% 
  head(n = 5)
```
:::

::::

:::: {style="display: grid; grid-template-columns: 1fr 1fr; grid-column-gap: 10px;"}


::: {}

SQL code: Get the release year and largest budget for all films, grouped by release year

```{sql, connection=con}
SELECT release_year, MAX(budget)
FROM films
GROUP BY release_year
LIMIT 5;
```

:::

::: {}
R code: Get the release year and largest budget for all films, grouped by release year

```{r warning=FALSE, message=FALSE}
films %>%
  group_by(release_year) %>% summarize(budget_max = max(budget)) %>%
  select(release_year, budget_max) %>% 
  head(n = 5)
```
:::

::::


:::: {style="display: grid; grid-template-columns: 1fr 1fr; grid-column-gap: 10px;"}


::: {}

SQL code: Get the country, release year, and lowest amount grossed per release year per country. Order your results by country and release year.

```{sql, connection=con}
SELECT country, release_year, MIN(gross)
FROM films
GROUP BY release_year, country
ORDER BY country, release_year
LIMIT 5;
```

:::

::: {}
R code: Get the country, release year, and lowest amount grossed per release year per country. Order your results by country and release year.

```{r warning=FALSE, message=FALSE}
films %>%
  group_by(release_year, country) %>% summarize(gross_min = min(gross)) %>%
  arrange(country, release_year) %>%
  select(country, release_year, gross_min) %>% 
  head(n = 5)
```
:::

::::




<!--
## Query 21: HAVING (REVISIT THIS)

- In SQL, aggregate functions can't be used in `WHERE` clauses.
- This means that you want to filter based on the result of an aggregate function, you need another way!


:::: {style="display: grid; grid-template-columns: 1fr 1fr; grid-column-gap: 10px;"}


::: {}

SQL code: the following query shows only those years in which > 10

```{sql, connection=con}
SELECT release_year
FROM films
GROUP BY release_year
HAVING COUNT(title) > 10;
--LIMIT 5;
```

:::

::: {}
R code: the following query shows only those years in which > 10
```{r warning=FALSE, message=FALSE}
#films %>%
#  count(title) %>%
#  group_by(release_year) %>% #count(title) %>% 
#  #mutate(n  = count(title)) %>%
#  filter(n > 10) %>%
#  select(release_year) #%>% 
#  #head(n = 5)
```
:::

::::
-->

<!--
## Query 15: `LIKE` and `NOT LIKE`

`WHERE` clause can be used to filter text data. However, so far you have been able to filter by specifying the exact text you have interested in. In the real world, often you will want to search for a pattern rather than a specific text string.

In SQL, the `LIKE` operator can be used in a `WHERE` clause to search for a pattern in a column. To accomplish this, you use something called a wildcard as a placeholder for some other values. There are two wildcards: `_` and `%`.

One can use `NOT LIKE` operator to find records that don't match the pattern you specify.

:::: {style="display: grid; grid-template-columns: 1fr 1fr; grid-column-gap: 10px;"}


::: {}

SQL code: Get the names of all people whose names begin with `B`. 

```{sql, connection=con}
SELECT name
FROM people
WHERE name LIKE 'B%'
LIMIT 5;
```

:::

::: {}
R code: Get the names of all people whose names begin with `B`. 

```{r warning=FALSE, message=FALSE}
people %>% 
  filter(name %like% "B%") %>% 
  select(name) %>%
  head(n = 5)
```
:::

::::
-->

# Joining Data

## Query 21: INNER JOINS

- `INNER JOIN` and `LEFT JOIN`  are probably the two most common joins.
- `id` field is called `key` since it is used to reference both the left and right tables
- `INNER JOIN` gathers the value columns from both the left and right tables with common `key` `ids` in both.

::::


:::: {style="display: grid; grid-template-columns: 1fr 1fr; grid-column-gap: 10px;"}


::: {}

SQL code: INNER JOIN and SELECTing all columns

```{sql, connection=con}
SELECT city.name AS city, country.region AS region
FROM cities AS city
  -- Inner join to countries
  INNER JOIN countries AS country
  -- Match on the country codes
  ON city.country_code = country.code
LIMIT 5;
```

:::

::: {}
R code: INNER JOIN and SELECTing all columns

```{r warning=FALSE, message=FALSE}
cities %>%
  inner_join(countries, by = c("country_code" = "code")) %>%
  select(city = name, region) %>%
  head(n = 5)
```
:::

::::


## Query 22: INNER JOINS - 2

::::


:::: {style="display: grid; grid-template-columns: 1fr 1fr; grid-column-gap: 10px;"}


::: {}

SQL code: 

```{sql, connection=con}
SELECT c.code AS country_code, year, inflation_rate
FROM countries AS c
  INNER JOIN economies AS e
  ON c.code = e.code
LIMIT 5;
```

:::

::: {}
R code: 

```{r warning=FALSE, message=FALSE}
countries %>%
  inner_join(economies, by = "code") %>%
  select(country_code = code, year, inflation_rate) %>%
  head(n = 5)
```
:::

::::


## Query 23: INNER JOIN via USING

::::


:::: {style="display: grid; grid-template-columns: 1fr 1fr; grid-column-gap: 10px;"}


::: {}

SQL code: 

```{sql, connection=con}
SELECT c.continent, l.name AS language, l.official
FROM countries AS c
  INNER JOIN languages AS l
  USING (code)
LIMIT 5;
```

:::

::: {}
R code: 

```{r warning=FALSE, message=FALSE}
countries %>%
  inner_join(languages, by = "code") %>%
  select(continent, language = name, official) %>%
  head(n = 5)
```
:::

::::

## Query 24: CASE WHEN and THEN

- Often it's useful to look at a numerical field not as raw data, but instead as being in different categories or groups.
- You can use `CASE` with `WHEN`, `THEN`, `ELSE`, and `END` to define a new grouping field.

::::



:::: {style="display: grid; grid-template-columns: 1fr 1fr; grid-column-gap: 10px;"}


::: {}

SQL code: 

```{sql, connection=con}
SELECT continent, code, surface_area,
    CASE WHEN surface_area > 2000000 THEN 'large'
        WHEN surface_area > 350000 AND surface_area < 2000000 THEN 'medium'
        ELSE 'small' END
        AS geosize_group
FROM countries
LIMIT 5;
```

:::

::: {}

R code: 

```{r warning=FALSE, message=FALSE}
countries %>%
  mutate(geosize_group = case_when(
    surface_area > 2000000 ~ "large",
    surface_area > 350000 & surface_area < 2000000 ~ "medium",
    surface_area < 350000 ~ "small"
  )) %>%
  select(continent, code, surface_area, geosize_group) %>%
  head(n = 5)
```
:::

::::

